The complex network of trophic interactions in a subAntarctic oceanic Marine Protected

### 2023-07-04 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
Schloss & Riccialdelli sug.

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, 53º–55ºS and 56º–62ºW, respectively), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterise the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1788 predator-prey interactions and 380 species, with a link density of 4.70 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although structural properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, such as the South Georgia and South Sandwich, Bouvet, Prince Edward, and Macquarie islands [@IUCN2023]. Interestingly, the case of the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, 53º–55ºS 56º–62ºW), which is the focus of this work, is unique since these MPAs are associated to a submarine plateau and its southern adjacent deep slope region, respectively [@Falabella2017]. In addition, such MPAs are part of a network of protected areas in the sub-Antarctic area (jointly with MPA Yaganes) that aims to protect and strengthen the functioning of this southern region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers (or proportions) of endemic species, and/or high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a sensitive and biodiverse benthic habitat that was barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by high biomass of vulnerable and fragile species (mainly Porifera and Cnidaria) that meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResourcesCCAMLR2009]. Also, it provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge and nursery area [@GarciaAlonso2018; @Vazquez2018; @Delpiani2020; @Troccoli2020; @CovattiAle2022; @Fischer2022; @Florencia2023]. The maintenance of this singular community is related to local and regional oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020]. Overall, 811 benthic and plankton species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time in the area in recent years [@AdministraciondeParquesNacionales2022]. However, identifying the main species involved in the maintenance of ecosystem services and health as well as for management and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem are regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure, meaning a shorter food-chain length and a greater trophic overlap and redundancy, than other sub-Antarctic areas, such as the continental shelf off Tierra del Fuego. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].

High-latitude marine ecosystems, such as the MPA N-BB, are considered complex systems in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @AdministraciondeParquesNacionales2022], a better understanding of species interactions' complexity and structure is needed.
This aspect can be tackled by analysing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats to the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022]; also fishing vessels are allowed to operate in the western section of the MPA N-BB (i.e. Marine National Reserve category), altering the stocks of commercially important fish species [@Martinez2021; @AdministraciondeParquesNacionales2022]. Moreover, there is a potential hazard related to the effects of offshore activities (exploration and explotation) to the west of the MPA N-BB [@AdministraciondeParquesNacionales2022].

In the present work, we present the first detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this, we applied a network approach to a highly resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and identify the species' role in such a network framework.

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws 26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB II, respectively [@Tombesi2020; @AdministraciondeParquesNacionales2022] (Figure 1). They are located 150 km east of Isla de los Estados and 200 km south of Malvinas/Falkland Islands.
The MPA N-BB I comprises nearly 34,000 km^2^ circumscribed by the 200 m isobath, between 54º–55ºS and 56º–62ºW, with a slight slope extended nearly 370 km east–west. Physical features in the BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC [@Guerrero1999; @Acha2004; @Piola2009]. The BB is surrounded by steep flanks of up to 4000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. The N-BB II includes such a deep slope, protecting about 32,000 km^2^. Intense upwelling and mixing occur over it, entraining deep nutrient-rich waters into the photic layer [@Piola2009; @Matano2019] and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019].

@@

In order to build the network of predator-prey interactions, we reviewed more than 150 references considering published articles, Ph.D. theses, public databases, and reports belonging to 16 research cruises conducted in the MPAs N-BB I and II during 2014-2019. It is noteworthy that the sampling effort was greater in the MPA N-BB I. Furthermore, we took into account personal communications from experts belonging to the working group of the study area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the authors' expertise contributing to the present study was a key factor in enhancing the quality of the network, and inherently improved the network representation. A list of the references used to build the network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the study area, we followed the concept of trophic species, here defined as aggregated groups of taxa. In most cases, we followed it when specific data on species, in the taxonomic sense, were not available. In some cases, we collapsed species when taxa shared the same set of predators and prey (trophic similarity, @Martinez1991), one of the aggregation methods that better preserve food web functional properties [@Gauzens2013]. In addition, for endemic species (e.g. bryozoan *Burdwoodipora paguricola*) and other species with no trophic studies so far, we inferred their feeding interactions applying a conservative approach that assumes that the set of prey and predators are at some point preserved in time. In those cases we gathered information from upper taxonomic levels (i.e. Genus, Family, Order, Class, Phylum) as a good proxy variable [@Morales-Castilla2015; @Pomeranz2019]. Details about this can be found in Supplementary Material (Table S2).

@@

The species-level analysis aims to describe the species' role in the food web. For this, we considered the following properties: (1) betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4) topological role TR; and (5) trophic level TL (Table 1). Topological roles refer to the fact that food webs tend to naturally organize in non-random, modular patterns, where modules are defined as a group of species that interact more frequently among themselves than with species that are not members of the module [@Guimera2005]. Species can play different roles in this respect, according to the pattern of interactions within their own module and/or across modules. We computed the topological role for each species, classified as module hub, species with a relatively high number of interactions, but most within its own module; module specialist, species with relatively few interactions and most within its own module; module connector, species with relatively few interactions mainly between modules; and network connector, species with high connectivity between and within modules [@Guimera2005].

@@

In terms of complexity, the MPA Namuncurá - Burdwood Bank food web (N-BB I and II) consisted of 1788 predator-prey interactions and 380 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density (e.g., the average number of interactions per species) of 4.70, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient higher than the random networks (Table 2).

@@

The degree distribution of the food web showed an asymmetric frequency in the number of interactions, where most of the species had a relatively low number of interactions and few species concentrated most of them (Figure 3A). The distribution of prey among predators showed that most consumers fed on a low number of prey whereas few had multiple prey (Figure 3B). The top-five predators in number of prey were: yellowfin notothen *Patagonotothen guntheri* (Notothenioid fish, 50 prey), rock cod *Patagonotothen ramsayi* (Notothenioid fish, 49 prey), broad nose skate *Bathyraja brachyurops* (Chondrichthyan, 33 prey), Patagonian toothfish *Dissostichus eleginoides* (Notothenioid fish, 30 prey), and graytail skate *Bathyraja griseocauda* (Chondrichthyan, 28 prey). Following the same distribution pattern, few prey presented multiple predators (Figure 3C). The top-five prey in number of predators were: Detritus (Non-living, 153 predators), the three categories of Diatoms considered (benthic, centric and pennate, 72.5 predators on average), and species of the genus *Euphausia* (Zooplankton, 46 predators). Finally, taking into account the interactions within each functional group, most interactions were concentrated in a few species (Figure 3D). The most evident species were: *Doryteuthis gahi* (Cephalopoda), *Grimothea gregaria* (Decapoda), *Patagonotothen ramsayi*, *Patagonotothen guntheri* and *Dissostichus eleginoides* (bentho-pelagic fish), *Sprattus fuegensis* and *Micromesistius australis* (pelagic fish), and species of *Euphausia* and *Themisto gaudichaudii* (Zooplankton).
Overall, there is an evident asymmetry in the distribution of interactions among species at different levels in the MPA N-BB food web.

@@

The food web of the MPA N-BB ecosystem analysed in this study is one of the most highly-resolved networks of trophic interactions ever studied, not only for a high-latitude marine ecosystem but also for any marine protected area worldwide to our knowledge. It is of paramount importance to consider the complexity of species interactions in order to gain insights into the structure and functioning of the ecosystem, since the aggregation of species might mask food web properties and produce type II errors (false positive) [@Martinez1993; @Gauzens2013].

Food web connectance is a feature that resumes the complexity of the network, but more importantly, it is an emergent property of pairwise species interactions [@Poisot2014]. It contains information regarding how interactions within an ecological network are distributed and predicts reasonably well key dynamical properties of ecological networks [@Dunne2002a]. Complex marine food webs (i.e. with more than 25 trophic species) show connectance values ranging from 0.01 - 0.27 [@Marina2018a]. In particular, food webs from high-latitude regions tend to exhibit a connectance closer to the minimum (between 0.01 and 0.05) [@Santana2013; @Kortsch2015; @Rodriguez2022]. Whether food webs display a low or a high connectance helps to better comprehend ecosystem's synthetic properties like robustness. In this sense, empirical analyses support the notion that highly-connected ecological networks are robust against external perturbations such as the introduction of new (e.g., invasive) species [@Smith-Ramesh2017] as well as species removal (e.g., local extinction) [@Dunne2002; @Montoya2003]. The connectance of the food web of the MPA Namuncurá - Burdwood Bank (0.01) is one of the lowest reported so far for these regions; in particular, it appears to much lower than that of Beagle Channel (0.05), an adjacent coastal area [@Rodriguez2022].

@@

It is suggested that the small-world pattern, i.e., a network with short path length and high clustering coefficient, is not frequent in complex marine food webs, mainly due to a low clustering coefficient compared to random networks [@Dunne2002b; @Marina2018a]. However, the food web of the MPA N-BB does display a small-world pattern. Consequences of this could be of great importance in recognizing evolutionary paths and the vulnerability to perturbations [@Montoya2002]. On the one hand, a short path length implies a rapid spread of an impact (e.g., contaminant, population fluctuation, local extinction) throughout the network but, at the same time, more potentially adaptive dynamics in the face of external perturbations [@Montoya2002; @Williams2002]. On the other hand, a high clustering coefficient indicates the formation of subnetworks composed only by the neighbours of particular species. This translates into a greater resistance of the network due to the confinement of perturbations mainly within subnetworks and not spreading between them [@Kortsch2019; @Heer2020]. Overall, a small-world topology provides ecological networks with greater resilience and resistance [@Bornatowski2017; @Dormann2017].

@@

In summary, the food web of the MPA N-BB presents a combination of network properties that makes it unique in terms of network resolution, complexity, and structural pattern. All this suggests that the ecosystem might be fragile to external perturbations targeting highly connected species, which in turn coincides to be commercial exploited species as fishes [@Martinez2015]. However, structural properties might provide resilience and resistance with the final outcome of a rearranged structure maintaining its functions.

@@

The degree distribution allows identifying important species, such as potential keystone species [@Sole2001; @Dunne2002], generalist/specialist species, and dominant food sources [@Kondoh2010].

We have identified that most of the consumers in the food web of the MPA N-BB either have a narrow diet or are specialists, while few present a broad diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Belleggia2008a], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4), acting as predator and prey, they might be important links between lower and higher trophic levels. This result is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed in the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources from which most consumers depend on and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016], likely feeding on detritus that is constantly resuspended from the bottom [@Martin2021]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators such as the black-browed and grey-headed albatrosses (*Thalassarche melanophris* and *Thalassarche chrysostoma*, respectively) [@Catry2004], and baleen whales (species of the genera Balaenoptera and Eubalaena) [@Valenzuela2018].

@@

We compiled information on the species and trophic diversity of the Marine Protected Area Namuncurá - Burdwood Bank, generating an unprecedented, well-resolved network of trophic interactions for a sub-Antarctic ecosystem, identifying the complexity and structure of the system, and the main species role in a network framework. Particular properties at the network level allowed us to identify the ecosystem's vulnerability and potential response to perturbations in the presence of highly-connected species, with a rearranged structure maintaining their functions due to its potential resilience and resistance.

>>>>

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, 53º–55ºS and 56º–62ºW, respectively), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterise the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1788 predator-prey interactions and 380 species, with a link density of 4.70 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although other properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, such as the South Georgia and South Sandwich, Bouvet, Prince Edward, and Macquarie islands [@IUCN2023]. Interestingly, the case of the MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, 53º–55ºS 56º–62ºW), which is the focus of this work, is unique since these MPAs are associated to a submarine plateau and its southern adjacent deep slope region, respectively [@Falabella2017]. In addition, such MPAs are part of a network of protected areas in the sub-Antarctic area (jointly with MPA Yaganes) that aims to protect this southern region in order to contribute to global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers (or proportions) of endemic species, and/or high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a potentially sensitive and biodiverse benthic habitat that was only barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by high biomass of vulnerable and fragile species (mainly Porifera, Bryozoa and Cnidaria) that considered with their environment meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResourcesCCAMLR2009]. Also, the benthic realm provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge and nursery area [@GarciaAlonso2018; @Vazquez2018; @Delpiani2020; @Troccoli2020; @CovattiAle2022; @Fischer2022; @Matusevich2022; @Florencia2023]. The maintenance of this singular community is related to local and regional oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020].

Overall, 811 benthic and plankton species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time in the area in recent years [@AdministraciondeParquesNacionales2022]. Identifying the main species involved in the maintenance of ecosystem services and health as well as for management and conservation is essential. Recently, the structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem are regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure, meaning a shorter food chain and a greater trophic overlap and redundancy, than other sub-Antarctic areas, such as the continental shelf off Tierra del Fuego. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].

High-latitude marine ecosystems, such as the MPA N-BB, are complex systems in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @AdministraciondeParquesNacionales2022], a better understanding of species interactions' complexity and structure is needed.
This aspect can be tackled by analysing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats to the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022]; also fishing vessels are allowed to operate in the western section of the MPA N-BB (i.e. Marine National Reserve category), altering the stocks of commercially important fish species [@Martinez2021; @AdministraciondeParquesNacionales2022]. Moreover, there is a potential hazard related to the effects of offshore activities (exploration and explotation) to the west of the MPA N-BB [@AdministraciondeParquesNacionales2022].

In the present work, we present the first detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this, we applied a network approach to a highly resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and identify the species' role in the network.

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws 26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB II, respectively [@Tombesi2020; @AdministraciondeParquesNacionales2022] (Figure 1). They are located 150 km east of Isla de los Estados and 200 km south of Malvinas/Falkland Islands.
The MPA N-BB I comprises nearly 34,000 km^2^ circumscribed by the 200 m isobath, between 54º–55ºS and 56º–62ºW, with a slight slope extended nearly 370 km east–west. Physical features in the BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC [@Guerrero1999; @Acha2004; @Piola2009]. The BB is surrounded by steep flanks of up to 4000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. The N-BB II includes such a deep slope, protecting about 32,000 km^2^. Intense upwelling and mixing occur in relation with the slope, entraining deep nutrient-rich waters into the photic layer [@Piola2009; @Matano2019] and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019].

@@

In order to build the network of predator-prey interactions, we reviewed more than 170 references considering published articles, Ph.D. theses, public databases, and reports belonging to 16 research cruises conducted in the MPAs N-BB I and II during 2014-2019. It is noteworthy that the sampling effort was greater in the MPA N-BB I. Furthermore, we took into account personal communications from experts belonging to the working group of the study area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the authors' expertise contributing to the present study was a key factor in enhancing the quality of the network, and inherently improved the network representation. A list of the references used to build the network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the study area, we followed the concept of trophic species, here defined as follows: taxa collapsed into a single node in the network. In most cases, we followed this concept when specific data on species, in the taxonomic sense, were not available. In some cases, we collapsed species when taxa shared the same set of predators and prey (trophic similarity, @Martinez1991), one of the aggregation methods that better preserve food web functional properties [@Gauzens2013]. In addition, for endemic species (e.g. bryozoan *Burdwoodipora paguricola*) and other species with no trophic studies so far, we inferred their feeding interactions applying a conservative approach that assumes that the set of prey and predators are at some point preserved in time. In those cases we gathered information from upper taxonomic levels (i.e. Genus, Family, Order, Class, Phylum) as a good proxy variable [@Morales-Castilla2015; @Pomeranz2019]. Details about this can be found in Supplementary Material (Table S2). Furthermore, we considered non-living food sources, such as detritus and necromass, as prey species in the food web context.

@@

The species-level analysis aims to describe the species' role in the food web. For this, we considered the following properties: betweenness Btw, closeness Cl, trophic similarity TS, topological role TR, and trophic level TL (Table 1). Topological roles refer to the fact that food webs tend to naturally organize in non-random, modular patterns, where modules are defined as a group of species that interact more frequently among themselves than with species that are not members of the module [@Guimera2005]. Species can play different roles in this respect, according to the pattern of interactions within their own module and/or across modules. We computed the topological role for each species, classified as module hub, species with a relatively high number of interactions, but most within its own module; module specialist, species with relatively few interactions and most within its own module; module connector, species with relatively few interactions mainly between modules; and network connector, species with high connectivity between and within modules [@Guimera2005].

@@

In terms of complexity, the MPA Namuncurá - Burdwood Bank food web consisted of 1788 predator-prey interactions and 380 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density (e.g., the average number of interactions per species) of 4.70, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient higher than the random networks (Table 2).

@@

The degree distribution of the food web showed an asymmetric frequency in the number of interactions, where most of the species had a relatively low number of interactions and few species concentrated most of them (Figure 3A). The distribution of prey among predators showed that most consumers fed on a low number of prey whereas few had multiple prey (Figure 3B). The top-five predators in number of prey were: yellowfin notothen *Patagonotothen guntheri* (Notothenioid fish, 50 prey), rock cod *Patagonotothen ramsayi* (Notothenioid fish, 49 prey), broad nose skate *Bathyraja brachyurops* (Chondrichthyan, 33 prey), Patagonian toothfish *Dissostichus eleginoides* (Notothenioid fish, 30 prey), and graytail skate *Bathyraja griseocauda* (Chondrichthyan, 28 prey). Following the same distribution pattern, few prey presented multiple predators (Figure 3C). The top-five prey (or food sources) in number of predators were: Detritus (Non-living, 153 predators), the three categories of Diatoms considered (benthic, centric and pennate, 72.5 predators on average), and species of the genus *Euphausia* (Zooplankton, 46 predators). Finally, taking into account the interactions within each functional group, most interactions were concentrated in a few species (Figure 3D). The most evident species were: *Doryteuthis gahi* (Cephalopoda), *Grimothea (=Munida) gregaria* (Decapoda), *Patagonotothen ramsayi*, *Patagonotothen guntheri* and *Dissostichus eleginoides* (bentho-pelagic fish), *Sprattus fuegensis* and *Micromesistius australis* (pelagic fish), and species of *Euphausia* and *Themisto gaudichaudii* (Zooplankton).
Overall, there is an evident asymmetry in the distribution of interactions among species at different levels in the MPA N-BB food web.

@@

The food web of the MPA N-BB ecosystem analysed in this study is one of the most highly-resolved networks of trophic interactions ever studied, not only for a high-latitude open-ocean ecosystem but also for any marine protected area worldwide to our knowledge. It is of paramount importance to consider the complexity of species interactions in order to gain insights into the structure and functioning of the ecosystem, since the aggregation of species might mask food web properties and produce type II errors (false positives) [@Martinez1993; @Gauzens2013].

Food web connectance is a feature that resumes the complexity of the network, but more importantly, it is an emergent property of pairwise species interactions [@Poisot2014]. It contains information regarding how interactions within an ecological network are distributed and predicts reasonably well key dynamical properties of ecological networks [@Dunne2002a]. Complex marine food webs (i.e. with more than 25 trophic species) show connectance values ranging from 0.01 - 0.27 [@Marina2018a]. In particular, food webs from high-latitude regions tend to exhibit a connectance closer to the minimum (between 0.01 and 0.05) [@Santana2013; @Kortsch2015; @Rodriguez2022]. Whether food webs display a low or a high connectance helps to better comprehend ecosystem's synthetic properties like robustness. In this sense, empirical analyses support the notion that highly-connected ecological networks are robust against external perturbations such as the introduction of new (e.g., invasive) species [@Smith-Ramesh2017] as well as species removal (e.g., local extinction) [@Dunne2002; @Montoya2003]. The connectance of the food web of the MPA Namuncurá - Burdwood Bank (0.01) is one of the lowest reported so far for these regions; in particular, it appears to be much lower than that of Beagle Channel (0.05), an adjacent coastal area [@Rodriguez2022].

@@

It is suggested that the small-world pattern, i.e., a network with short path length and high clustering coefficient, is not frequent in complex marine food webs, mainly due to a low clustering coefficient compared to random networks [@Dunne2002b; @Marina2018a]. However, the food web of the MPA N-BB does display a small-world pattern. Consequences of this could be of great importance in recognizing species evolutionary paths and the vulnerability to perturbations [@Montoya2002]. On the one hand, a short path length implies a rapid spread of an impact (e.g., contaminant, population fluctuation, local extinction) throughout the network but, at the same time, more potentially adaptive dynamics in the face of external perturbations [@Montoya2002; @Williams2002]. On the other hand, a high clustering coefficient indicates the formation of subnetworks composed only by the neighbours of particular species. This translates into a greater resistance of the network due to the confinement of perturbations mainly within subnetworks and not spreading between them [@Kortsch2019; @Heer2020]. Overall, a small-world topology provides ecological networks with greater resilience and resistance [@Bornatowski2017; @Dormann2017].

@@

In summary, the food web of the MPA N-BB presents a combination of network properties that makes it unique in terms of network resolution, complexity, and structural pattern. All this suggests that the food web might be fragile to external perturbations targeting highly connected species, which in turn coincides to be commercial exploited species as fishes [@Laptikhovsky2013; @Martinez2015; @Winter2023]. However, structural properties might provide resilience and resistance with the final outcome of a rearranged structure maintaining its functions.

@@

The degree distribution allows identifying important species, such as potential keystone species (i.e. highly connected) [@Sole2001; @Dunne2002], generalist/specialist species, and dominant food sources [@Kondoh2010].

We have identified that most of the consumers in the food web of the MPA N-BB either have a narrow diet or are specialists, while few present a broad or generalist diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Belleggia2008a], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4), acting as predator and prey, they might be important links between lower and higher trophic levels. This result is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed in the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources on which most consumers depend and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016], likely feeding on detritus that is constantly resuspended from the bottom [@Martin2021]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators such as the black-browed and grey-headed albatrosses (*Thalassarche melanophris* and *Thalassarche chrysostoma*, respectively) [@Catry2004], and baleen whales (species of the genera Balaenoptera and Eubalaena) [@Valenzuela2018].

@@

We compiled information on the species and trophic diversity of the oceanic Marine Protected Area Namuncurá - Burdwood Bank, generating an unprecedented, well-resolved network of trophic interactions for a sub-Antarctic ecosystem, identifying the complexity and structure of the system, and the main species role in a network framework. Particular properties at the network level allowed us to identify the ecosystem's vulnerability and potential response to perturbations in the presence of highly-connected species, with a rearranged structure maintaining their functions due to its potential resilience and resistance.
